---
title: "Homework 6"
author: "Laya Buchanan"

date: 2020-19-05
output: github_document
---

This is my submission for the sixth homework assignment for P8105.  

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Problem 1

Read in the data.

```{r}


birthweight_df = 
  read_csv("./data/birthweight.csv") %>% 
  mutate(babysex = recode(babysex, '1' = "male", '2' = "female"),
         malform = recode(malform, '0' = "absent", '1' = "present"),
         mrace = recode(mrace, '1' = "white", '2' = "black", '3' = "asian", '4' = "puerto rican", '8' = "other"),
         frace = recode(frace, '1' = "white", '2' = "black", '3' = "asian", '4' = "puerto rican", '8' = "other", '9' = "unknown")
         )
```
#### Model Building

I'm going to the variable babysex to see if it has an apparent crude association with baby birth weight using a simple plot. I will also be looking at the linearity of the association.

```{r}
birthweight_df %>% 
  ggplot(aes(x = bwt, y = babysex)) + 
  geom_point()
```


In this case, we see an association with a binomial dependent variable, so it will be appropriate to build a linear model to examine this association. Next, I'm going to build a model to assess the relationship between the sex of the baby and baby's birth weight.  I will not be adding any other variables to this model because none of the other variables are potential confounders.


```{r}
babysex_fit = lm(bwt ~ babysex, data = birthweight_df)
```

Next, I create a plot of model residuals against fitted values.

```{r}
birthweight_df %>% 
  add_residuals(babysex_fit) %>% 
  add_predictions(babysex_fit) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point()
```

Now, I will create two other models to compare my proposed model to, one sing length at birth and gestational age as predictors (main effects only), and the other using head circumference, length, sex, and all interactions (including the three-way interaction) between these.

First, I will assess the linearity of these relationships.

```{r}
birthweight_df %>% 
  ggplot(aes(x = bwt, y = gaweeks)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = bwt, y = gaweeks)) + 
  geom_point()
```

```{r}
gaweeks_fit = gam(bwt ~ gaweeks, data = birthweight_df)

interactions_fit = gam(bwt ~ babysex + bhead + blength + babysex*bhead + babysex*blength + blength*bhead + babysex*bhead*blength, data = birthweight_df)
```

Next, I will make this comparison in terms of the cross-validated prediction error.

```{r}
cv_df = 
  crossv_mc(birthweight_df, 250) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) 
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(
    babysex_fit = map(train, ~lm(bwt ~ babysex, data = birthweight_df)),
    gaweeks_fit = map(train, ~gam(bwt ~ parity + gaweeks, data = birthweight_df)),
    interactions_fit = map(train, ~gam(bwt ~ babysex + bhead + blength + babysex*bhead + babysex*blength + blength*bhead + babysex*bhead*blength, data = birthweight_df))) %>% 
  mutate(
    rmse_babysex = map2_dbl(babysex_fit, test, ~rmse(model = .x, data = .y)),
    rmse_gaweeks = map2_dbl(gaweeks_fit, test, ~rmse(model = .x, data = .y)),
    rmse_interactions = map2_dbl(interactions_fit, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Based on the distributions of the root mean squared errors, the model featuring head circumference, length, sex, and all interactions between these has the best fit, while the model I proposed has the worst fit.

## Problem 3

First, import the weather data.
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
I am creating 5000 bootstrap samples from the data imported above, for each bootstrap sample, I am producing estimates of r squared and log(β0*β1) and then plotting their distribution.
```{r}
boot_straps = 
  weather_df %>% 
  bootstrap(100, id = "strap_number")

log_dist = 
  boot_straps %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results) %>% 
  select(strap_number, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>%
  janitor::clean_names() %>%  
  mutate(
    b0b1 = log(intercept*tmin)
  )

log_dist %>% 
ggplot(aes(x = b0b1)) + geom_density() +
  labs(
    title = "log(β0*β1 Distribution")

r_dist = 
  boot_straps %>% 
  mutate(
    models = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::glance)) %>% 
  select(results) %>% 
  unnest(results)

r_dist %>% 
  ggplot(aes(x = r.squared)) + geom_density() +
  labs(
    title = "r Squared Distribution")
```

They both follow a roughly normal distribution.

#### Confidence Intervals

Next, I will identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1).

```{r}
log_dist %>% 
  summarize(
    ci_lower = quantile(b0b1, 0.025), 
    ci_upper = quantile(b0b1, 0.975)) %>% 
  knitr::kable(digits = 3)

r_dist %>% 
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975)) %>% 
  knitr::kable(digits = 3)
```



